---
title: "HW 06 Reproducible Report"
author: "Jacob Kebe"
date: "2024-10-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



## Data Visualization

```{r pressure, echo=FALSE}
leadData <- read.csv("~/Documents/GitHub/Kebe_Project_01/DataProcessed/lead-iq-01_processed.csv")
boxplot(IQ~Smelter, data = leadData)
```

Here we used a boxplot to visualize the IQ's for the "Near" group and the "Far Group". The Far group's IQ visually seems to be significantly better than the Near group. \
## Means by Group
```{r means}
avg_iq_near <- mean(leadData[leadData$Smelter == "Near",]$IQ)
avg_iq_near

avg_iq_far <- mean(leadData[leadData$Smelter == "Far",]$IQ)
avg_iq_far
```
Here we have our mean IQs for the two groups, "Near" and "Far" respectively.

## Table

```{r table, echo = FALSE}
library(knitr)
#dfSummary(leadData)
#summary stats near iq
near_iq_stats <- summary(leadData[leadData$Smelter=="Near",]$IQ)
#summary stats far iq
far_iq_stats <- summary(leadData[leadData$Smelter=="Far",]$IQ)

#create dataframe and kable it
stat_names <- c("Min.","1st Qu.", "Median","Mean","3rd Qu.", "Max")
iq_stats <- data.frame(nearlab =stat_names, Near = as.numeric(near_iq_stats), farlab = stat_names, Far = as.numeric(far_iq_stats))

kable(iq_stats, col.names= c("","IQ: Near Group", "", "IQ: Far Group"), digits = 2)
```
This table provides summary statistics for our two groups' IQs. We have reported the minimum, maximum, quantiles, and measures of center. The Far group's mean and median and quantiles are higher, but the minimum is smaller than the near group. \


**Code used for plots and analysis**

``` r
#load data
leadData <- read.csv("~/Documents/GitHub/Kebe_Project_01/DataProcessed/lead-iq-01_processed.csv")

#Calculate averages
avg_iq_near <- mean(leadData[leadData$Smelter == "Near",]$IQ)
avg_iq_near

avg_iq_far <- mean(leadData[leadData$Smelter == "Far",]$IQ)
avg_iq_far

#Create plot
boxplot(IQ~Smelter, data = leadData)



library(knitr)
#summary stats near iq
near_iq_stats <- summary(leadData[leadData$Smelter=="Near",]$IQ)
#summary stats far iq
far_iq_stats <- summary(leadData[leadData$Smelter=="Far",]$IQ)

#create dataframe and kable it
stat_names <- c("Min.","1st Qu.", "Median","Mean","3rd Qu.", "Max")
iq_stats <- data.frame(nearlab =stat_names, Near = as.numeric(near_iq_stats), farlab = stat_names, Far = as.numeric(far_iq_stats))

kable(iq_stats, col.names= c("","IQ: Near Group", "", "IQ: Far Group"), digits = 2)
```
